---
title: "Quick Start with the `difNLR()` function"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start with the `difNLR()` function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,       # show code
  eval = TRUE,       # run code
  results = 'markup',# show printed output
  warning = TRUE,    # show warnings
  message = TRUE,    # show messages
  fig.width = 6,     # default figure width
  fig.height = 4,    # default figure height
  fig.align = "center"
)
```

This vignette demonstrates the **basic usage of the `difNLR()` function** using example data from the package.

## Load package

```{r loading package}
library(difNLR)
```

## Load data

We here use the `GMAT` dataset. It is s a generated dataset based on parameters
from Graduate Management Admission Test (GMAT, Kingston et al., 1985). First two
items were considered to function differently in uniform and non-uniform way
respectively. The dataset represents responses of 2,000 subjects to
multiple-choice test of 20 items. A correct answer is coded as 1 and incorrect
answer as 0. The column group represents group membership, where 0 indicates
reference group and 1 indicates focal group. Groups are the same size (i.e.
1,000 per group). The distributions of total scores (sum of correct answers) are
the same for both reference and focal group (Martinkova et al., 2017).

```{r loading data}
data(GMAT)
Data <- GMAT[, 1:20] # binary items
group <- GMAT[, "group"] # group membership variable

summary(Data)
table(group)
```

## DIF detection with generalized logistic regression models

Not constrained 4 parameter logistic (PL) model using the IRT parametrization is
of the following form:
$$
\begin{align}
P(Y_{pi} = 1 \mid X_p, G_p) =& (c_{i} + c_{i\text{DIF}} \cdot G_p) + (d_{i} + d_{i\text{DIF}} \cdot G_p - c_{i} - c_{i\text{DIF}} \cdot G_p) / \\
 &(1 + \exp(-(a_i + a_{i\text{DIF}} \cdot G_p) \cdot (X_p - b_p - b_{i\text{DIF}} \cdot G_p))),
\end{align}
$$
where \(X_p\) is the matching criterion (e.g., standardized total score) and
\(G_p\) is a group membership variable for respondent \(p\). Parameters
\(a_i\), \(b_i\), \(c_i\), and \(d_i\) are discrimination, difficulty,
guessing, and inattention for the reference group for item \(i\). Terms
\(a_{i\text{DIF}}\), \(b_{i\text{DIF}}\), \(c_{i\text{DIF}}\), and
\(d_{i\text{DIF}}\) then represent differences between the focal and reference
groups in discrimination, difficulty, guessing, and inattention for item
\(i\), respectively.

To perform DIF detection, besides `Data` and `group` variable, we need to
specify the name of the focal group with the argument `focal.name` and the
generalized regression model through the `model` argument. The options are as
follows: \code{Rasch} for 1PL model with discrimination parameter fixed on value
1 for both groups, \code{1PL} for 1PL model with discrimination parameter set
the same for both groups, \code{2PL} for logistic regression model, \code{3PLcg}
for 3PL model with fixed guessing for both groups, \code{3PLdg} for 3PL model
with fixed inattention for both groups, \code{3PLc} (alternatively also
\code{3PL}) for 3PL regression model with guessing parameter, \code{3PLd} for
3PL model with inattention parameter, \code{4PLcgdg} for 4PL model with fixed
guessing and inattention parameter for both groups, \code{4PLcgd} (alternatively
also \code{4PLd}) for 4PL model with fixed guessing for both groups,
\code{4PLcdg} (alternatively also \code{4PLc}) for 4PL model with fixed
inattention for both groups, or \code{4PL} for 4PL model.

Here we use the 3PL model with the same guessing parameter \(c_i\) for both
groups on the `GMAT` dataset.
```{r difNLR with 3PL model}
(x <- difNLR(Data, group, focal.name = 1, model = "3PLcg"))
```
The functions computes \(\chi^2\)-statistics and corresponding p-values, suggesting
items 1, 2, 7, and 19 to function differently for the reference and focal groups. 

## Plot item characteristic curves

We plot item characteristic curves for DIF items.
```{r plot difNLR}
plot(x, item = x$DIFitems)
```
